Spfy: an integrated graph-based sequence database for real-time prediction of Escherichia coli phenotypes using the SuperPhy platform

### 2018-02-22 paper-webserver.tex
Intro edits.

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}. In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to replicate the results of traditional wet-lab tests; this allows new WGS results to be viewed in the context of historical tests.

@@

% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
Traditionally, these platforms have been applied to samples numbering only a few dozen while WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years. For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.

Many of the comparative analyses that are run are broadly useful, and therefore computed multiple times. To begin to compute results on all available genomes, an effective method to mitigate the recomputing of analyses is to make the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology. Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.

@@

% 5. how we solved it
In this study, we present the Spfy platform, which integrates a graph database with real-time analysis options. By integrating analysis options with data storage, we avoid recomputing analysis for identical runs. Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis modules as they are developed. The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

>>>>

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the identification of known antimircrobial (AMR) resistance genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, {ResFinder}, {ARG-ANNOT}, and {ARIBA}; the identification of known virulence genes through software such as VirulenceFinder \cite{kleinheinz2014applying}, {SRST2}, and {an additional reference}; the annotation of bacterial genomes, through software such as Prokka \cite{doi:10.1093/bioinformatics/btu153}, {another}, and {another}. \todo{Add a section on serotyping and subtyping a la Phylotyper}. These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.

@@

% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
Traditionally, these platforms have been applied to samples numbering only a few dozen \cite{} while WGS of bacterial pathogens have recently accumulated in public databases in the hundreds of thousands, with millions set to be available within the next few years. For \textit{Escherichia coli} alone, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million sequenced genomes in GenBank \cite{doi:10.1093/nar/gks1195}.

Many of the comparative analyses that are currently used in the analyses of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes. An effective method to mitigate the recomputing of analyses, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology. Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.

@@

% 5. how we solved it
In this study, we present the Spfy update to the SuperPhy platform, which integrates a graph database with real-time analyses; this integration avoids recomputing identical analyses. Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed. The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
